Uncertainty Quantification in Multiscale Models of Charge Transport in Organic Semiconductors: Influence of the Exhange-Correlation Functional

### 2025-01-20 paper.tex
base draft of discussion

\begin{abstract}
  \bjoern{TODO}{rewrite}This work quantifies uncertainties in a multiscale model for charge dynamics in organic semiconductors (OSCs). A first-principle multiscale model integrates classical molecular dynamics (MD), density functional theory (DFT), quantum mechanical calculation and continuous-time random walk (CTRW) processes to allow the simulation of charge transport in OSCs. The model brings inherent uncertainties arising from the empirical approximations and numerical simulations. Our work particularly focusing on quantifying the uncertainty in DFT exchange-correlation functional.

  By investigating the effects of varying Hartree-Fock (HF) levels within DFT functionals, we analyze the impact on electronic structures such as reorganization energy, molecule energy distribution, and coupling elements. The primary goal is to assess the robustness of predicting the quantity of interest (QoI), that is time-of-flight and charge mobility, under these uncertainties.

  Monte Carlo simulations and sensitivity analysis are employed to estimate the range and confidence levels of the predicted charge mobility. As shown by Sobol indices, the QoI is most sensitive to molecule energy, then reorganization energy and least sensitive to coupling elements.  
\end{abstract}

@@

\bjoern{STOP}{HERE}
\begin{itemize}
\item \ahfx, while only a proxy for the use of different functionals in MMM, gives a good indication the challenges in obtaining quantutative predictions of charge mobilities
\item uncertainty for a low/moderate disorder material covering a range of one order of magnitude 
\item field-dependence much less affected
\item uncertainty mostly stems from site energy uncertainties, in a broader sense this highlights the importance for accurate methods to determine these energies not just as a distribution
\end{itemize}

>>>>

\begin{abstract}
This study investigates the impact of exchange-correlation functional choices on the predictive accuracy of multiscale models for charge transport in organic semiconductors (OSCs). A hybrid functional approach is applied to analyze uncertainties in key parameters influencing charge mobility, focusing on the Hartree--Fock exchange fraction . Using 2-methyl-9,10-bis(naphthalen-2-yl)anthracene (MADN) as a test system, molecular dynamics and density functional theory are combined to compute reorganization energies, site energies, and electronic coupling elements. Monte Carlo sampling quantifies the uncertainty propagation, revealing that site energy variations dominate transport property uncertainties, while coupling elements contribute minimally. The findings underscore the need for accurate parameter determination and functional selection, with implications for enhancing the reliability of first-principles-based multiscale modeling frameworks in OSC design.  
  %\bjoern{TODO}{rewrite}This work quantifies uncertainties in a multiscale model for charge dynamics in organic semiconductors (OSCs). A first-principle multiscale model integrates classical molecular dynamics (MD), density functional theory (DFT), quantum mechanical calculation and continuous-time random walk (CTRW) processes to allow the simulation of charge transport in OSCs. The model brings inherent uncertainties arising from the empirical approximations and numerical simulations. Our work particularly focusing on quantifying the uncertainty in DFT exchange-correlation functional.

%  By investigating the effects of varying Hartree-Fock (HF) levels within DFT functionals, we analyze the impact on electronic structures such as reorganization energy, molecule energy distribution, and coupling elements. The primary goal is to assess the robustness of predicting the quantity of interest (QoI), that is time-of-flight and charge mobility, under these uncertainties.

%  Monte Carlo simulations and sensitivity analysis are employed to estimate the range and confidence levels of the predicted charge mobility. As shown by Sobol indices, the QoI is most sensitive to molecule energy, then reorganization energy and least sensitive to coupling elements.  
\end{abstract}

@@

\bjoern{STOP}{HERE}
\section{Discussion}
The findings presented in this study provide valuable insights into the sensitivity and uncertainty of multiscale models for charge transport in organic semiconductors (OSCs) due to the choice of exchange-correlation functionals in density functional theory (DFT). By systematically varying the Hartree--Fock exchange fraction  in hybrid functionals, the study uncovers critical dependencies and sources of uncertainty in the computational prediction of charge mobility.

One key observation is the dominance of site energy uncertainties in influencing the variability of the time-of-flight (ToF) and mobility predictions. This reflects the substantial impact of electrostatic and polarization interactions, which are integral to site energy determination, on charge transport dynamics. In contrast, the uncertainties in reorganization energies and electronic coupling elements are shown to have a relatively minor effect. This suggests that further efforts to enhance the accuracy of site energy calculations—potentially through improved parameterization or advanced quantum mechanical methods—could significantly improve the reliability of multiscale modeling predictions.

Interestingly, the results also demonstrate that the uncertainties in electronic coupling elements, particularly those with very small magnitudes, are less relevant for overall charge transport. This aligns with the percolation analysis, which indicates that coupling elements below a critical threshold contribute minimally to the connectivity of the charge transport network. Thus, simplifying approximations in handling these low-value couplings could be justified in specific scenarios to reduce computational complexity without compromising accuracy.

Another notable finding is the robustness of field-dependent mobility predictions. Despite uncertainties in the underlying molecular parameters, the electric field dependence of mobility exhibits relatively consistent Poole--Frenkel behavior across the tested range of \ahfx values. This robustness suggests that multiscale models retain predictive value for field-dependent trends, even when parameter uncertainties are present.

However, the study also highlights the challenge of achieving quantitative accuracy in charge mobility predictions, given that the uncertainty range spans an order of magnitude. This raises important considerations for interpreting simulation results in the context of experimental data. For instance, small differences in predicted mobility between two materials or designs may not be statistically significant given the intrinsic uncertainties.

The use of Monte Carlo sampling and Sobol sensitivity analysis in this study provides a powerful framework for systematically quantifying the contributions of different sources of uncertainty to model predictions. The results clearly demonstrate that site energies are the dominant source of variability, with a Sobol index of 0.95. This indicates that future developments in multiscale modeling should prioritize reducing uncertainties in site energy calculations to enhance the reliability of predicted charge transport properties.

Despite the progress made, there are limitations to this study that warrant further investigation. First, the analysis is restricted to variations in \ahfx as a proxy for different exchange-correlation functionals. While this approach captures a significant portion of the uncertainty, other factors such as basis set selection, dispersion corrections, and model approximations may also contribute to variability in the results. Expanding the uncertainty quantification framework to include these factors would provide a more comprehensive understanding of the robustness of multiscale models.

Furthermore, the study focuses on a single material system, MADN, which exhibits moderate disorder. Extending the analysis to other OSCs with varying degrees of disorder, charge-carrier types, or different molecular morphologies would enhance the generality of the conclusions.

\section{Conclusion and Future Directions}

In conclusion, this study underscores the importance of uncertainty quantification in multiscale modeling of OSCs. The dominant role of site energy uncertainties highlights the need for accurate quantum mechanical methods and parameterization techniques in this area. Additionally, the robustness of field-dependent mobility trends suggests that multiscale models remain valuable tools for guiding material design, even in the presence of parameter uncertainties.

Future research should aim to expand the scope of uncertainty quantification to include additional sources of variability, such as alternative functionals and morphological complexities. Integrating machine learning approaches to refine parameter estimates and leveraging high-throughput simulations could further enhance the predictive power of multiscale models in the design of next-generation OSCs.

The results for the quantification of the uncertainty from functional choices in charge transport properties obtained from first-principles based multiscale modeling for the disordered system of MADN have revealed

\begin{itemize}
\item \ahfx, while only a proxy for the use of different functionals in MMM, gives a good indication the challenges in obtaining quantutative predictions of charge mobilities
\item uncertainty for a low/moderate disorder material covering a range of one order of magnitude 
\item field-dependence much less affected
\item uncertainty mostly stems from site energy uncertainties, in a broader sense this highlights the importance for accurate methods to determine these energies not just as a distribution
\end{itemize}
